---
title: "Meeting Notes"
date: "Updated: `r format(Sys.Date())`"
output: 
  html_document:
    highlight: pygments
    theme: yeti
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

suppressPackageStartupMessages({
  library(tidyverse);
  library(here);
})

```

# Meeting 2020-04-14 w Mike

## To Do: 

 - discuss updating the sims with diff distributions
    - focus on Poisson, Gaussian, and uniform
 - review paper Caren sent
 - write up manuscript without field experiments for now? 
 - update and submit the regulation paper

## eDNA manuscript

**make a conceptual model:** and use to set up intro and paper. Check Bioscience paper that Sarah wrote for guidance.

 - schematic of different distributions with variables labeled
 
### Distributions

```{r}
# make a pois distrib:
set.seed(111)
pois <- rpois(100, lambda = 3)
binom <- rbinom(100, 20, 0.2)
nbinom <- rnbinom(100, 20, 0.2)
norm <- rnorm(100)

y<-seq(1:100)
plot(y, dbinom(y, 100, 0.1), type = "l", col="salmon", ylim = c(0,.15))
lines(y, dnbinom(y, 100, .9), col="slateblue", ylab="")
lines(y, dnorm(y, mean=50, sd = 4), col="gray", lwd=2)

```


**Main question is** how to identify the number of haplos present that we see across multiple loci.  Parameters that control this are both biological and technological. 

 - how much variation? (*biological*)
 - how many individuals are present in a sample (*biological*)?
 - how much coverage? if 1x coverage, even if you have 1000 haplos may not pick up anything (*technological*)
 
**Coverage and Length Matter Most**

 - coverage and length matter, number of loci doesn't matter as much
 - No of individuals in a sample will effect number of haplos in sample
 - Amount of diversity (relates to read length, more reads means more potential haplos)
 - Coverage will pick up more or less haplos

**Technical vs biological issues** 

 - readout, distribution shifts with variation among indivs, but also with tech issues (coverage, loci length)
 number of haplos vs. number of loci
 - that distrigution is effected by technological and biological
 - variation in number of haplos in a sample is biological

## Review: 

 - $\mu$ = mutation rate, we assumed general mutation rate of 1e^-8^
 - N~e~ = Coalescent effective size, using conservative est. of 25,000
 - conclusion 3: I think our upward loci length of 10,000 was more of a "future" estimate. We are close with single read options but we prob aren't there yet. 
 - thinking about $\theta$ = 4N$\mu$ can be misleading because N~e~ is coalescent effective population size which is usually orders of magnitude bigger than local census size. For many species our highest theta (10) would be more like a 2 kb DNA fragment. Mean read lengths of 5+ kb are available now with PacBio, but the output needs to increase to make it less cost prohibitive. The limitation will become DNA fragment length in eDNA samples at some point, if not already, even as sequencing tech improves. 